### 2020-06-08 manuscript/manuscript.tex

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[english,,man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdftitle={Analytic reproducibility in articles receiving open data badges at Psychological Science: An observational assessment},
            pdfkeywords={open data, badges, reproducibility, open science, meta-research, journal policy},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\makeatother
\shorttitle{Analytic reproducibility}
\author{Tom E. Hardwicke\textsuperscript{1,2}, Manuel Bohn\textsuperscript{3}, Kyle MacDonald\textsuperscript{4}, Emily Hembacher\textsuperscript{3}, Michèle B. Nuijten\textsuperscript{5}, Benjamin N. Peloquin\textsuperscript{3}, Benny deMayo\textsuperscript{3}, Bria Long\textsuperscript{3}, Erica J. Yoon\textsuperscript{3}, \& Michael C. Frank\textsuperscript{3}}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} University of Amsterdam\\\textsuperscript{2} Charité – Universitätsmedizin Berlin\\\textsuperscript{3} Stanford University\\\textsuperscript{4} University of California, Los Angeles\\\textsuperscript{5} Tilburg University}
\authornote{Tom E. Hardwicke, Department of Psychology, University of Amsterdam \& Meta-Research Innovation Center Berlin (METRIC-B), QUEST Center for Transforming Biomedical Research, Charité – Universitätsmedizin Berlin. Manuel Bohn, Emily Hembacher, Benjamin N. Peloquin, Benny deMayo, Bria Long, Erica J. Yoon, Michael C. Frank, Department of Psychology, Stanford University. Kyle MacDonald, Department of Communication, University of California, Los Angeles. Michèle B. Nuijten, Department of Methodology and Statistics, Tilburg School of Social and Behavioral Sciences, Tilburg University.

Correspondence concerning this article should be addressed to Tom E. Hardwicke, Nieuwe Achtergracht 129B, Department of Psychology, University of Amsterdam, 1018 WT Amsterdam, The Netherlands. E-mail: tom.hardwicke@uva.nl}
\note{}
\keywords{open data, badges, reproducibility, open science, meta-research, journal policy}
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=english]{babel}
\else
  % load polyglossia as late as possible as it *could* call bidi if RTL lang (e.g. Hebrew or Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\fi

\title{Analytic reproducibility in articles receiving open data badges at Psychological Science: An observational assessment}

\date{}

\abstract{
For any scientific report, repeating the original analyses upon the original data should yield the original outcomes. We evaluated analytic reproducibility in 25 Psychological Science articles awarded open data badges. Initially, 16 (64\%, CI {[}43,81{]}) articles contained at least one ``major numerical discrepancy'' (\textgreater{}10\% difference) prompting us to request input from original authors. Ultimately, target values were reproducible without author involvement for 9 (36\%, CI {[}20, 59{]}) articles; reproducible with author involvement for 6 (24\%, CI {[}8, 47{]}) articles; not fully reproducible with no substantive author response for 3 (12\%, CI {[}0, 35{]}) articles; and not fully reproducible despite author involvement for 7 (28\%, CI {[}12, 51{]}) articles. Overall, 37 major numerical discrepancies remained out of 789 checked values 5\% CI {[}3,6{]}, but original conclusions did not appear affected. Non-reproducibility was primarily caused by unclear analysis reporting. Open data alone is not sufficient to ensure analytic reproducibility.
}

\begin{document}
\maketitle

\begin{verbatim}
## Warning: Duplicated column names deduplicated: 'Quote of statement, if provided'
## => 'Quote of statement, if provided_1' [27], 'Page number of statement, if
## provided' => 'Page number of statement, if provided_1' [28]
\end{verbatim}

A minimum quality standard expected of all scientific manuscripts is that any numerical values can be reproduced if the original analyses are repeated upon the original data (Bollen et al., 2015). This concept is known as \emph{analytic reproducibility} (LeBel et al., 2018; or relatedly, \emph{computational reproducibility}\footnote{Assessments of computational reproducibility typically attempt to reproduce values by re-running original computational code with the original data and can therefore fail if original code is unavailable or non-functioning {[}e.g., @stodden2018{]}. By contrast, assessments of analytic reproducibility typically attempt to reproduce values by repeating the original analysis procedures, which can involve implementing those procedures in new code {[}e.g., @hardwicke2018a{]}.}, Stodden et al., 2018). When a number cannot be reproduced, this minimally indicates that the process by which it was calculated has not been sufficiently documented. Non-reproducibility may also indicate that an error has occurred, either during the original calculation or subsequent reporting. Either way, there is a breakdown in the integrity of the analysis pipeline that transforms raw data into reported results. As a result, non-reproducibility can create uncertainty about the provenance and veracity of scientific evidence, potentially undermining the credibility of any associated inferences. As a result, non-reproducibility can undermine data re-use (Hardwicke, Mathur, et al., 2018), complicate replication attempts (Nuijten et al., 2018), and create uncertainty about the provenance and veracity of scientific evidence, potentially undermining the credibility of any associated inferences (LeBel et al., 2018).

Difficulty establishing the analytic reproducibility of research reports has been encountered in several scientific domains, including economics, political science, and psychology (Chang \& Li, 2015; Eubank, 2016; Hardwicke, Mathur, et al., 2018; Obels et al., 2019; Stodden et al., 2018). A preliminary obstacle for many such studies is that research data are typically unavailable (Hardwicke, Thibault, et al., 2020; Hardwicke \& Ioannidis, 2018; Wicherts et al., 2006). Even when data can be accessed, suboptimal management and inadequate documentation can complicate data reuse (Hardwicke, Mathur, et al., 2018; Kidwell et al., 2016). These failures to adequately preserve and share earlier stages of the research pipeline typically preclude downstream assessment of analytic reproducibility.

A previous study in the domain of psychology largely circumvented data availability issues by capitalizing on a mandatory open data policy introduced at the journal Cognition (Hardwicke, Mathur, et al., 2018). After assessing a sample of 35 articles with available data that had already passed initial quality checks, at least one value could not be reproduced within a 10\% margin of error in 24 of the 35 (69\%) articles. Reproducibility issues were resolved in 11 of the 24 articles after consultation with original authors yielded additional data or clarified analysis procedures. Ultimately, 64 of 1324 (5\%) checked values could not be reproduced despite author involvement. Importantly, there were no clear indications that the reproducibility issues seriously undermined the conclusions of the original articles. Nevertheless, this study highlighted a number of quality control and research transparency issues, including suboptimal data management; unclear, incomplete, or incorrect analysis specification; and reporting errors.

In the present study, we intended to investigate whether the findings of Hardwicke, Mathur, et al. (2018) extended to a corpus of Psychological Science articles that received an \enquote{Open Data Badge} to signal data availability. In order to focus specifically on the reproducibility of the analysis process rather than upstream data availability or management practices, we selected only articles that had reusable data according to a previous study (Kidwell et al., 2016). The submission guidelines of Psychological Science specifically stated that authors could earn an Open Data badge for \enquote{making publicly available the digitally-shareable data necessary to reproduce the reported result.}\footnote{See \url{https://perma.cc/SFV8-DAZ6} (originally retrieved 9th October, 2017).} Additionally, authors were asked to self-certify that they had provided \enquote{\ldots{}sufficient information for an independent researcher to reproduce the reported results.}\footnote{See \url{https://perma.cc/N8K7-DXP9?type=image} (originally retrieved 9th October, 2017).} Thus, if this policy was operating as intended, all numbers presented in these articles should be independently reproducible. If not, we hoped to learn about causes of non-reproducibility in order to identify areas for improvement in journal policy and research practice.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

The study protocol (hypotheses, methods, and analysis plan) was pre-registered on October 18th, 2017 (\url{https://osf.io/2cnkq/}). All deviations from this protocol or additional exploratory analyses are explicitly acknowledged in Supplementary Information \ref{prereg}.

\hypertarget{design}{%
\subsection{Design}\label{design}}

We employed a retrospective non-comparative case-study design based on Hardwicke, Mathur, et al. (2018). The numerical discrepancy between \enquote{old} values reported in the target articles and \enquote{new} values obtained in our analyses was quantified using percentage error\footnote{An important caveat of this measure should be noted: large percentage differences can occur for small absolute differences when the absolute magnitude of values is small (and vice versa). Thus, when considering individual cases, quantitative measures should be evaluated in the full context of our case-by-case qualitative observations (Supplementary Information \ref{vignettes}).} (\(PE = \frac{|new-old|}{old}\times100\)). Numerical discrepancies were classified as \enquote{minor} (0\% \textgreater{} PE \textless{} 10\%) or \enquote{major} (PE \textgreater{}= 10\%). If an old \emph{p} value fell on the opposite side of the \(\alpha = .05\) boundary relative to a new \emph{p} value we additionally recorded a \enquote{decision error}. We recorded if there was insufficient information to complete aspects of the analyses.

We classified articles as \enquote{not fully reproducible} (any major numerical discrepancies, decision errors, or insufficient information to proceed) or \enquote{reproducible} (no numerical discrepancies or minor numerical discrepancies only) and noted whether author involvement was provided or not. We recorded the potential causal loci of non-reproducibility and judged the likelihood that non-reproducibility impacted conclusions drawn in the original articles. Team members provided a subjective estimate of the time they spent on each reproducibility check. Additional procedural details are available in Supplementary Information @ref(sup\_design).

\hypertarget{sample}{%
\subsection{Sample}\label{sample}}

The sample was based on a corpus of psychology articles that had been examined in a previous project investigating the impact of an \enquote{open badges} scheme introduced at Psychological Science (Kidwell et al., 2016). This sample was selected because the open badges scheme and assessment of reusability by Kidwell et al. (2016) enabled us to largely circumvent upstream issues related to data availability and data management and instead focus on downstream issues related to analytic reproducibility. A precision analysis indicates that the sample size affords adequate precision for the purposes of gauging policy compliance (Supplementary Information @ref(sup\_sample)).

Of 47 articles marked with an open data badge, Kidwell and colleagues had identified 35 with datasets that met four reusability criteria (accessible, correct, complete, and understandable). For each of these articles, one investigator (TEH) attempted to identify a coherent set of descriptive and inferential statistics, roughly 2-3 paragraphs of text, sometimes including a table or figure, related to a \enquote{substantive} finding based on \enquote{relatively straightforward} analyses. In total, 789 discrete numerical values reported in 25 articles were identified as target values. The articles were published between January 2014 and May 2015. Further information about the sample is available in Supplementary Information @ref(sup\_sample).

\hypertarget{procedure}{%
\subsection{Procedure}\label{procedure}}

The primary aim of the reproducibility checks was to recover the reported target values by repeating the original analysis (as described in the original articles and any additional documentation such as supplementary materials, codebook, analysis scripts) upon the available data. Attempting alternative analyses was outside the scope of the study. To minimize error and facilitate reproducibility, each reproducibility check was conducted by at least two team members who documented the reanalysis process in an R Markdown report (available at \url{https://osf.io/hf9jy/}). If articles were initially classified as not fully reproducible, we emailed the author(s) of the article and used any additional information they provided to try and resolve the issues before a final classification was determined. Additional procedural details are available in Supplementary Information @ref(sup\_design).

\hypertarget{data-analysis}{%
\subsection{Data analysis}\label{data-analysis}}

The results can be considered at several layers of granularity. Detailed qualitative information about each reproducibility check is available in the individual reproducibility reports (\url{https://osf.io/hf9jy/}) and summarized in a short \enquote{vignette} available in Supplementary Information \ref{vignettes}. We report descriptive and inferential statistics at the article-level and value-level. We report descriptive and inferential\footnote{Note that although one goal of the study was to characterize analytic reproducibility within a finite sample as a follow-up to Kidwell et al. (2016), we were also interested in generalizing to other psychology articles. To address this second goal, we use standard inferential statistics, recognizing that, because they are generated from a convenience sample, the generality of our findings may be limited.} statistics at the article-level and value-level. Ninety-five per cent confidence intervals (CIs) are displayed in square brackets.

\hypertarget{results}{%
\section{Results}\label{results}}

Prior to author involvement, our reproducibility checks indicated that 16 out of the 25 (64\%, CI {[}43,81{]}) articles contained at least one major numerical discrepancy. After requesting input from original authors, several issues were resolved through the provision of additional information\footnote{All except one case involved the provision of additional information that was not included in the original article or clarification of original information that was ambiguous or incomplete.In the exception case (4-1-2015\_PS), the authors pointed out that we had missed a relevant footnote in the original article.} leaving 10 out of the 25 (40\%, CI {[}22,61{]}) articles containing at least one major numerical discrepancy. Even after author contact, we could not complete 3 (12\%, CI {[}3,32{]}) reproducibility checks because insufficient information was available about aspects of the original analyses.

Ultimately, target values were fully reproducible for 9 (36\%, CI {[}20, 59{]}) articles without requiring author involvement; fully reproducible for 6 (24\%, CI {[}8, 47{]}) articles only with author involvement; not fully reproducible for 7 (28\%, CI {[}12, 51{]}) articles despite author involvement; and not fully reproducible for 3 (12\%, CI {[}0, 35{]}) articles with no substantive response from authors. Availability of original analysis scripts appeared to provide some modest benefits in terms of reproducibility outcomes and time expenditure (see Supplementary Information @ref(sup\_results)).

In no cases did the observed numerical discrepancies appear to be consequential for the conclusions stated in the original articles (see Supplementary Information \ref{vignettes}).

\begin{figure}
\centering
\includegraphics{figs/scatterPlot-1.pdf}
\caption{\label{fig:scatterPlot}Scatterplot showing reported p-values as a function of reanalysis p-values, classified by reproducibility outcome. Axes are on a log scale. For display purposes, 41 values below 0.001 are not shown.}
\end{figure}

After author involvement, 37 major numerical discrepancies remained amongst the 789 target values examined across all articles (5\% CI {[}3,6{]}). This included 1 decision error for which we obtained a statically significant p value of 0.02 in contrast to the reported non-significant p-value of 0.09. A scatterplot illustrating the consistency of old and new p-values is displayed in Figure \ref{fig:scatterPlot}. The frequency of numerical discrepancies by value types is displayed in Figure \ref{fig:valueTypePlot}.

\begin{figure}
\centering
\includegraphics{figs/valueTypePlot-1.pdf}
\caption{\label{fig:valueTypePlot}Frequency of reproducibility outcomes by value type. Variation/uncertainty measures include standard deviations, standard errors, and confidence intervals. Effect sizes include Cohen's d, Person's r, partial eta squared, and phi. Test statistics include t, F, and chi squared. Central tendency measures include means and medians.}
\end{figure}

Where possible we attempted to identify the causal locus of the reproducibility issues we encountered though this was not always possible to confirm definitively. Supplementary Figure \ref{fig:locusPlot} shows the frequency of four types of discrete causal loci that we determined were involved in non-reproducibility and how many of these issues were resolved through author involvement. The most common issues we encountered were related to unclear, incomplete, or incorrect reporting of analytic procedures. Examples include unidentified statistical tests, unclear aggregation procedures, non-standard p-value reporting, and unreported data exclusions. Most of these issues could be resolved when original authors provided additional information. Less frequently, we encountered typographical errors and some issues related to data files, including erroneous or missing data. For many instances of non-reproducibility, the causal locus remained unidentified even after contact with original authors.

Team members provided concurrent estimates of their time spent on each stage of the analysis. Altogether, they estimated that they spent between 1 and 30 (median = 7, interquartile range = 5) hours actively working on each reproducibility check (Supplementary Figure \ref{fig:timePlot}; total time = 213 hours). This estimate excludes time spent waiting on internal (within our team) and external (with original authors) communications.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

A reasonable expectation of any scientific manuscript is that repeating the original analyses upon the original data will yield the same quantitative outcomes (Bollen et al., 2015). We have found that this standard of analytic reproducibility was frequently not met in a sample of Psychological Science articles receiving open data badges. Importantly, none of the reproducibility issues we encountered appeared seriously consequential for the conclusions stated in the original articles. Nevertheless, non-reproducibility highlighted fundamental quality control and documentation failures during data management, data analysis, and reporting. Furthermore, the extensive time investment required to check and establish analytic reproducibility would likely be prohibitive for many researchers interested in re-using data. The findings are consistent with a previous study which found a similar rate of non-reproducibility and identified unclear, incomplete, or incorrect analysis reporting as a primary contributing factor (Hardwicke, Mathur, et al., 2018). Although the open badges scheme introduced at Psychological Science has been associated with an increase in data availability (Kidwell et al., 2016), the current findings suggest that additional efforts may be required in order to ensure analytic reproducibility.
Although some reproducibility issues were resolved through involvement of original authors, relying on such assistance is not ideal for several reasons: (1) it substantially increases the workload and time investment, both for researchers attempting to re-use data and for original authors; (2) important details about and materials related to the study are more likely to be forgotten or misplaced over time, reducing the ability of authors to assist (Vines et al., 2014); (3) authors may be unwilling or unable to respond, as was the case for three of the articles we examined. Author involvement highlighted that non-reproducibility was often caused by unclear, incomplete, or incorrect reporting of analytic procedures in the original papers, issues which could often be resolved through the provision of additional information or clarifications beyond what was previously reported. In other cases, authors informed us of errors in shared data files or analysis scripts, suggesting that at some stage the original data, analyses, and research report had become decoupled. For many issues, neither we nor the original authors could reproduce the target values and the causal locus of non-reproducibility remains unidentified; it was no longer possible to reconstruct the original analysis pipeline.

This study has a number of important limitations and caveats. Most pertinently, generalizing the observed reproducibility rates to other articles requires some caution as several characteristics of our sample are likely to have had a positive impact on the reproducibility rate. We specifically selected a sample of articles for which data were already available and screened against several reusability criteria (Kidwell et al., 2016). This is relevant because most psychology articles are not accompanied by raw data or analysis scripts at all (Hardwicke, Thibault, et al., 2020; Hardwicke \& Ioannidis, 2018; Wicherts et al., 2006), and even when data is available, suboptimal management and documentation can complicate re-use (Hardwicke, Mathur, et al., 2018). Thus, most psychology articles would likely fail a reproducibility check before reaching the stage of the analysis pipeline at which we began our assessments.

Additionally, all articles in the sample had been submitted to a leading psychology journal that had recently introduced a number of policy changes intended to improve research rigor and transparency (Eich, 2014). These included the open badges scheme, removing word limits for methods and results sections, and requesting explicit disclosure of relevant methodological information like exclusions. It is likely that these new initiatives positively impacted reproducibility, either through directly encouraging adoption of reproducible research practices or attracting researchers to the journal who were already inclined to use such practices. Some evidence, for example, implies that data availability is modestly associated with a lower prevalence of statistical reporting inconsistencies (Wicherts et al., 2011). In sum, several features of the current sample are likely to facilitate reproducibility relative to a more diverse population of psychology articles.

Whilst the current findings are concerning, non-reproducibility is fortunately a solvable problem, at least in principle. As noted above, a major barrier to analytic reproducibility is low availability of research resources like data and code (Hardwicke, Thibault, et al., 2020). Some evidence suggests that journal policies which mandate data sharing can be highly effective at increasing data availability (Hardwicke, Mathur, et al., 2018; Nuijten et al., 2017), apparently more so than voluntary badge schemes (Kidwell et al., 2016). Sharing of analysis scripts may help by providing veridical documentation of the analysis process which is not well captured in verbal prose (Hardwicke, Mathur, et al., 2018). However, as with data sharing, mere availability will not be sufficient to confer the potential benefits of analysis scripts. The utility of scripts depends on a range of factors, including clarity, structure, and documentation, as well the programming language that is used and whether it is in a proprietary format or requires special expertise to understand. In the present study, script availability appeared to offer modest benefits in terms of reproducibility outcomes and time expenditure; however, as only six articles shared scripts\footnote{Note that Psychological Science's author guidelines explicitly state that the criteria for an open data badge includes sharing \enquote{annotated copies of the code or syntax used for all exploratory and principal analyses.} See \url{https://perma.cc/SFV8-DAZ6} (originally retrieved 9th October, 2017).}, it is difficult to draw strong conclusions about their impact from this evidence alone.

Another potential target for policy intervention includes improving quality control at the journal level, potentially through independent assessment of analytic reproducibility; however, this would naturally require additional resources (for discussion see Sakaluk et al., 2014). Some journals, such as the American Journal of Political Science, pay for a third-party institute to verify the reproducibility of all manuscripts prior to publication (Jacoby et al., 2017). Ideally, initiatives intended to improve analytic reproducibility should undergo empirical scrutiny in order to evaluate costs and benefits and identify any policy shortfalls or unintended consequences (Hardwicke, Serghiou, et al., 2020).

Although journal policy is potentially helpful for incentivisation and verification, the reproducibility of a scientific manuscript is fundamentally within the control of its authors. Fortunately, a variety of tools are now available that allow for the writing of entirely reproducible research reports in which data, analysis code, and research reports are intrinsically linked (e.g., Vuorre \& Crump, 2020). Detailed guidance on data sharing, data management, and analytic reproducibility is available to support psychological scientists seeking to improve these practices in their own research (Klein et al., 2018). The present manuscript is an illustration of these practices in action (\url{https://doi.org/10.24433/CO.6557927.v1}). It should be noted that ensuring the reproducibility of a scientific manuscript requires a non-trivial time investment. Continued development of user-friendly tools that facilitate reproducibility and dedicated training may help to reduce this burden. Additionally, the costs of ensuring reproducibility could be offset by the benefits of improved workflow efficiency, error detection and mitigation, and potential for data re-use.

\hypertarget{conclusion}{%
\subsection{Conclusion}\label{conclusion}}

This study has highlighted that the analytic reproducibility of published psychology articles cannot be guaranteed. It is inevitable that some scientific manuscripts contain errors and imperfections, as researchers are only human and people make mistakes (Rouder et al., 2019). However, most of the issues we encountered in this study were entirely avoidable. Data availability alone is insufficient; further action is required to ensure the analytic reproducibility of scientific manuscripts.

\hypertarget{open-practices-statement}{%
\section{Open practices statement}\label{open-practices-statement}}

The study protocol (hypotheses, methods, and analysis plan) was pre-registered on October 18, 2017 (\url{https://osf.io/2cnkq/}). All deviations from this protocol or additional exploratory analyses are explicitly acknowledged in Supplementary Information \ref{prereg}. All data exclusions and measures conducted during this study are reported. All data (\url{https://osf.io/qzmpj/}), materials (\url{https://osf.io/awxt6/}), and analysis scripts (\url{https://osf.io/vwzn4/}) related to this study are publicly available on the Open Science Framework. To facilitate reproducibility this manuscript was written by interleaving regular prose and analysis code using knitr (Xie, 2015) and papaja (Aust \& Barth, 2020), and is available in a Code Ocean container (\url{https://doi.org/10.24433/CO.6557927.v1}) which re-creates the software environment in which the original analyses were performed. Analysis code, reports, and Code Ocean containers are also available for each reproducibility check (\url{https://osf.io/hf9jy/}).

\hypertarget{conflict-of-interest-statement}{%
\section{Conflict of interest statement}\label{conflict-of-interest-statement}}

All authors declare that there were no conflicts of interest.

\hypertarget{funding-statement}{%
\section{Funding statement}\label{funding-statement}}

This work received no specific funding. TEH's contribution was enabled by a general support grant awarded to the Meta-Research Innovation Center at Stanford (METRICS) from the Laura and John Arnold Foundation and a grant from the Einstein Foundation and Stiftung Charité awarded to the Meta-Research Innovation Center Berlin (METRIC-B).

\hypertarget{author-contributions}{%
\section{Author contributions}\label{author-contributions}}

TEH and MCF designed the study. TEH, MB, KM, EH, MBN, BNP, BdM, BL, EJY, and MCF conducted reproducibility checks. TEH performed the data analysis. TEH and MCF wrote the manuscript. X provided feedback on the manuscript. All authors gave final approval for publication.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

We are grateful to the authors of the original articles for their assistance with the reproducibility checks. We thank students from Stanford's Psych 251 class, who contributed to the initial reproducibility checks.

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-R-papaja}{}%
Aust, F., \& Barth, M. (2020). \emph{papaja: Create APA manuscripts with R Markdown}. \url{https://github.com/crsh/papaja}

\leavevmode\hypertarget{ref-bollen2015}{}%
Bollen, K., Cacioppo, J. T., Kaplan, R. M., Krosnick, J. A., \& Olds, J. L. (2015). \emph{Social, Behavioral, and Economic Sciences Perspectives on Robust and Reliable Science}. National Science Foundation.

\leavevmode\hypertarget{ref-chang2015}{}%
Chang, A. C., \& Li, P. (2015). \emph{Is Economics Research Replicable? Sixty Published Papers from Thirteen Journals Say "Usually Not"} (pp. 1--26). Board of Governors of the Federal Reserve System.

\leavevmode\hypertarget{ref-eich2014}{}%
Eich, E. (2014). Business Not as Usual. \emph{Psychological Science}, \emph{25}(1), 3--6. \url{https://doi.org/10.1177/0956797613512465}

\leavevmode\hypertarget{ref-eubank2016}{}%
Eubank, N. (2016). Lessons from a Decade of Replications at the Quarterly Journal of Political Science. \emph{PS: Political Science \& Politics}, \emph{49}(2), 273--276. \url{https://doi.org/10.1017/S1049096516000196}

\leavevmode\hypertarget{ref-hardwicke2018}{}%
Hardwicke, T. E., \& Ioannidis, J. P. A. (2018). Populating the Data Ark: An attempt to retrieve, preserve, and liberate data from the most highly-cited psychology and psychiatry articles. \emph{PLOS ONE}, \emph{13}(8), e0201856. \url{https://doi.org/10.1371/journal.pone.0201856}

\leavevmode\hypertarget{ref-hardwicke2018a}{}%
Hardwicke, T. E., Mathur, M. B., MacDonald, K., Nilsonne, G., Banks, G. C., Kidwell, M. C., Hofelich Mohr, A., Clayton, E., Yoon, E. J., Henry Tessler, M., Lenne, R. L., Altman, S., Long, B., \& Frank, M. C. (2018). Data availability, reusability, and analytic reproducibility: Evaluating the impact of a mandatory open data policy at the journal Cognition. \emph{Royal Society Open Science}, \emph{5}(8), 180448. \url{https://doi.org/10.1098/rsos.180448}

\leavevmode\hypertarget{ref-hardwicke2020b}{}%
Hardwicke, T. E., Serghiou, S., Janiaud, P., Danchev, V., Crüwell, S., Goodman, S. N., \& Ioannidis, J. P. (2020). Calibrating the Scientific Ecosystem Through Meta-Research. \emph{Annual Review of Statistics and Its Application}, \emph{7}(1), 11--37. \url{https://doi.org/10.1146/annurev-statistics-031219-041104}

\leavevmode\hypertarget{ref-hardwicke2020a}{}%
Hardwicke, T. E., Thibault, R. T., Kosie, J., Wallach, J. D., Kidwell, M. C., \& Ioannidis, J. (2020). \emph{Estimating the prevalence of transparency and reproducibility-related research practices in psychology (2014-2017)}. \url{https://doi.org/10.31222/osf.io/9sz2y}

\leavevmode\hypertarget{ref-jacoby2017}{}%
Jacoby, W. G., Lafferty-Hess, S., \& Christian, T. M. (2017). \emph{Should Journals Be Responsible for Reproducibility? \textbar{} Inside Higher Ed}. https://www.insidehighered.com/blogs/rethinking-research/should-journals-be-responsible-reproducibility.

\leavevmode\hypertarget{ref-kidwell2016}{}%
Kidwell, M. C., Lazarevi\a'c, L. B., Baranski, E., Hardwicke, T. E., Piechowski, S., Falkenberg, L.-S., Kennett, C., Slowik, A., Sonnleitner, C., Hess-Holden, C., Errington, T. M., Fiedler, S., \& Nosek, B. A. (2016). Badges to Acknowledge Open Practices: A Simple, Low-Cost, Effective Method for Increasing Transparency. \emph{PLOS Biology}, \emph{14}(5), e1002456. \url{https://doi.org/10.1371/journal.pbio.1002456}

\leavevmode\hypertarget{ref-klein2018}{}%
Klein, O., Hardwicke, T. E., Aust, F., Breuer, J., Danielsson, H., Mohr, A. H., Ijzerman, H., Nilsonne, G., Vanpaemel, W., \& Frank, M. C. (2018). A Practical Guide for Transparency in Psychological Science. \emph{Collabra: Psychology}, \emph{4}(1), 20. \url{https://doi.org/10.1525/collabra.158}

\leavevmode\hypertarget{ref-lebel2018}{}%
LeBel, E. P., McCarthy, R. J., Earp, B. D., Elson, M., \& Vanpaemel, W. (2018). A Unified Framework to Quantify the Credibility of Scientific Findings: \emph{Advances in Methods and Practices in Psychological Science}. \url{https://doi.org/10.1177/2515245918787489}

\leavevmode\hypertarget{ref-nuijten2018}{}%
Nuijten, M. B., Bakker, M., Maassen, E., \& Wicherts, J. M. (2018). Verify original results through reanalysis before replicating. \emph{Behavioral and Brain Sciences}, \emph{41}. \url{https://doi.org/10.1017/S0140525X18000791}

\leavevmode\hypertarget{ref-nuijten2017}{}%
Nuijten, M. B., Borghuis, J., Veldkamp, C. L. S., Dominguez-Alvarez, L., Van Assen, M. A. L. M., \& Wicherts, J. M. (2017). Journal Data Sharing Policies and Statistical Reporting Inconsistencies in Psychology. \emph{Collabra: Psychology}, \emph{3}(1), 31. \url{https://doi.org/10.1525/collabra.102}

\leavevmode\hypertarget{ref-obels2019}{}%
Obels, P., Lakens, D., Coles, N., Gottfried, J., \& Green, S. A. (2019). \emph{Analysis of Open Data and Computational Reproducibility in Registered Reports in Psychology}. \url{https://doi.org/10.31234/osf.io/fk8vh}

\leavevmode\hypertarget{ref-rouder2019}{}%
Rouder, J. N., Haaf, J. M., \& Snyder, H. K. (2019). Minimizing Mistakes in Psychological Science. \emph{Advances in Methods and Practices in Psychological Science}, \emph{2}(1), 3--11. \url{https://doi.org/10.1177/2515245918801915}

\leavevmode\hypertarget{ref-sakaluk2014}{}%
Sakaluk, J., Williams, A., \& Biernat, M. (2014). Analytic Review as a Solution to the Misreporting of Statistical Results in Psychological Science. \emph{Perspectives on Psychological Science}, \emph{9}(6), 652--660. \url{https://doi.org/10.1177/1745691614549257}

\leavevmode\hypertarget{ref-stodden2018}{}%
Stodden, V., Seiler, J., \& Ma, Z. (2018). An empirical analysis of journal policy effectiveness for computational reproducibility. \emph{Proceedings of the National Academy of Sciences}, \emph{115}(11), 2584--2589. \url{https://doi.org/10.1073/pnas.1708290115}

\leavevmode\hypertarget{ref-vines2014}{}%
Vines, T. H., Albert, A. Y. K., Andrew, R. L., D\a'ebarre, F., Bock, D. G., Franklin, M. T., Gilbert, K. J., Moore, J.-S., Renaut, S., \& Rennison, D. J. (2014). The Availability of Research Data Declines Rapidly with Article Age. \emph{Current Biology}, \emph{24}(1), 94--97. \url{https://doi.org/10.1016/j.cub.2013.11.014}

\leavevmode\hypertarget{ref-vuorre2020}{}%
Vuorre, M., \& Crump, M. (2020). \emph{Sharing and organizing research products as R packages}. \url{https://doi.org/10.31234/osf.io/jks2u}

\leavevmode\hypertarget{ref-wicherts2011}{}%
Wicherts, J. M., Bakker, M., \& Molenaar, D. (2011). Willingness to Share Research Data Is Related to the Strength of the Evidence and the Quality of Reporting of Statistical Results. \emph{PLoS ONE}, \emph{6}(11), e26828. \url{https://doi.org/10.1371/journal.pone.0026828}

\leavevmode\hypertarget{ref-wicherts2006}{}%
Wicherts, J. M., Borsboom, D., Kats, J., \& Molenaar, D. (2006). The poor availability of psychological research data for reanalysis. \emph{American Psychologist}, \emph{61}(7), 726--728. \url{https://doi.org/10.1037/0003-066X.61.7.726}

\leavevmode\hypertarget{ref-R-knitr}{}%
Xie, Y. (2015). \emph{Dynamic documents with R and knitr} (2nd ed.). Chapman; Hall/CRC. \url{https://yihui.org/knitr/}

\clearpage
\makeatletter
\efloat@restorefloats
\makeatother

\begin{appendix}
\hypertarget{sup_sample}{%
\section{Supplementary sample information}\label{sup_sample}}

We downloaded data from the Kidwell et al. (2016) study (available here:
https://osf.io/u6g7t/) and selected the 47 articles that had received an
`Open Data Badge'. Kidwell and colleagues had determined that 35 of
these articles were accompanied by data sets that were accessible,
correct, complete, and understandable. One team member (TEH) examined
each of these articles and attempted to identify a `substantive finding'
supported by a `relatively straightforward analysis' according to the
following operational definitions (adopted directly from Hardwicke et
al.~2018): (A) A `substantive finding' is one that supports a central
conclusion of the article under scrutiny. This is to some extent
subjective, but generally refers to findings supported in the article's
abstract and/or displayed in a figure/table; (B) `Relatively
straightforward analyses' involved behavioural data only and employed
quantitative techniques that would be commonly found in an introductory
level statistics textbook aimed at undergraduate psychology students
(e.g., {\textbf{???}}). Examples included means, medians, standard
deviations, confidence intervals, standardized effect sizes,
correlations, t-tests, and ANOVAs.

The first finding encountered in the article that met these eligibility
criteria was selected. The target values for the reproducibility checks
were a coherent set of descriptive and inferential statistics related to
this finding, typically 2-3 paragraphs of information, occasionally
accompanied by a table or figure. During this assessment, 9 articles
were excluded because they did not meet the eligibility criteria (ID
codes: 12-8-2014 PS, 21-7-2014 PS, 13-6-2014 PS, 18-7-2014 PS, 3-2-2015
PS, 6-3-2015 PS, 16-8-2014 PS, 19-4-2015 PS, 15-2-2015 PS). An
additional article was excluded because the data no longer appeared to
be available (contra Kidwell et al., 2016; ID code: 8-11-2014 PS). In
total, 25 eligible articles were identified. A bibliography file is
available where article ID codes can be matched to their references
(https://osf.io/rszey/). Note that there was a reporting error in our
pre-registered protocol which said that article 16-2-2015 was excluded
when it was not.

A precision analysis (not pre-registered and performed after study
completion) shows the range of expected margin of error (confidence
interval width) for the range of possible effect sizes given our sample
size of 25 articles (Figure \ref{fig:precisionCurve}). The maximum
margin of error is 0.20.

\begin{figure}
\centering
\includegraphics{figs/precisionCurve-1.pdf}
\caption{\label{fig:precisionCurve}Precision analysis yielding expected
margin of error (confidence interval width) relative to effect size
(proportion) for a sample size of 25.}
\end{figure}

\hypertarget{sup_design}{%
\section{Supplementary design and procedures}\label{sup_design}}

The most serious consequence of non-reproducibility is that it
undermines the credibility of associated scientific claims. We therefore
attempted to determine the extent to which any reproducibility issues we
encountered had substantial implications for related conclusions drawn
in the original papers. As noted in a previous investigation (Hardwicke
et al., 2018), this is a complex issue that cannot be straightforwardly
measured by a quantitative index. We therefore, considered this question
on a case by case basis, using multiple factors to inform our judgement,
including the presence/absence of decision errors, the quantity of and
type of target values that could not be reproduced, the difference in
magnitude of key values like effect sizes, and the specificity of the
hypothesis under scrutiny. This was a necessarily subjective exercise;
however the judgement for each reproducibility check was agreed upon by
the (two or more) team members who worked on it in addition to the first
author (TEH) and senior author (MCF). The rationale underlying each
judgement is outlined in the reproducibility vignettes in Supplementary
Materials \ref{vignettes}.

Determining the causal locus of non-reproducibility was not
straightforward because we often did not have access to comprehensive
information about the original analysis process. Additionally, there is
not necessarily a direct mapping between discrete causal loci and
reproducibility issues; multiple numerical discrepancies could be
attributable to a single cause or multiple causes. Sometimes the causal
locus of non-reproducibility became apparent after discussions with the
original authors, however, all classifications were made independently
by our team.

Comparisons between values were partly automated through the use of a
custom R function (https://osf.io/teuzb/) that accounted for rounding
errors by rounding new values to the same number of decimal places as
old values prior to computing the percentage error and classifying the
value as a match, minor discrepancy, or major discrepancy. Note that for
212 values reported relative to a threshold (e.g., p \textless{} 0.05)
or embedded in figures, we could not directly calculate percentage error
and it was necessary to `eyeball' (i.e., visually compare side-by-side)
the values to confirm a match.

All team members had started or completed graduate level training in
psychology and had experience conducting the types of data analyses
typically found in an introductory psychology statistics textbook (e.g.,
Field et al., 2012).

In the first contact emails we sent to original authors, we explained
the project, the issues we had encountered with their article, and
provided a copy of an interim reproducibility report. If we were
notified that the email was not delivered, we spent up to 5 minutes
searching for a more recent contact email address. If the corresponding
author(s) did not respond after two weeks, a reminder email was sent.

\hypertarget{sup_results}{%
\section{Supplementary results}\label{sup_results}}

Ninety-five per cent confidence intervals (CIs) displayed in square
brackets are based on the Wilson method with continuity correction for
binomial proportions ({\textbf{???}}) and the Sison--Glaz method for
multinomial proportions ({\textbf{???}}).

\begin{figure}
\centering
\includegraphics{figs/locusPlot-1.pdf}
\caption{\label{fig:locusPlot}Frequency of discrete causal loci underlying
non-reproducibility across all reproducibility checks. Note that some
articles contained multiple causal loci.}
\end{figure}

\begin{figure}
\centering
\includegraphics{figs/timePlot-1.pdf}
\caption{\label{fig:timePlot}Estimated number of active work hours spent by
our team on each reproducibility check, presented separately by
reproducibility outcome.}
\end{figure}

Six articles were accompanied by analysis code, including one with
Matlab scripts and SPSS syntax (8-12-2014\_PS), one with Matlab scripts
only (2-2-2015\_PS), two with SPSS syntax only (2-10-2014\_PS,
16-9-2014\_PS), one with an R script (8-5-2015\_PS), and one with SAS
syntax (16-2-2015\_PS). Three articles were reproducible without author
involvement; one article was reproducible with author involvement; one
article was not reproducible with author involvement; and one article
was not reproducible with no author response. Team members estimated
that they spent between 3 and 10 (median = 6.50, interquartile range =
3) hours actively working on each of the reproducibility checks for
articles with scripts.

\hypertarget{prereg}{%
\section{Deviations from pre-registered protocol}\label{prereg}}

The pre-registered protocol (https://osf.io/2cnkq/) does not state that
we will report confidence intervals. However, we decided to report
confidence intervals as this would aid inferences beyond the sample.

The pre-registered protocol used the terminology ``major/minor errors''.
To improve clarity, we use ``major/minor numerical discrepancies''.

The pre-registered protocol used the terminology ``reproducibility
failure'' and ``reproducibility success'' in reference to article
classifications. To improve clarity, we use ``not fully reproducible''
and ``reproducible''.

The pre-registered protocol stated that ``After contacting authors for
assistance, a maximum time-limit of 2 months for resolution of any
issues will be observed. Any reproducibility issues that cannot be
resolved within this time-period will be considered reproducibility
failures for the purposes of the present investigation.'' We decided not
to impose this time limit as it seemed unreasonable that delays on our
side could influence reproducibility outcomes. Authors were able to
respond until the project was completed on June 2, 2020.

\hypertarget{vignettes}{%
\section{Reproducibility vignettes}\label{vignettes}}

Each of the reproducibility checks is summarised below in a short
vignette. Each vignette contains a link to a full R Markdown report
rendered in HTML that will take the reader through the reanalysis
process step-by-step. A link is also provided to an OSF/Github
repository that contains all relevant data and analysis code and a link
to a Code Ocean container which recreates the software environment in
which the original analyses were run to facilitate reproducibility.

\hypertarget{vignette-1-article-2-2-2015_ps}{%
\subsection{Vignette 1 (article
2-2-2015\_PS)}\label{vignette-1-article-2-2-2015_ps}}

Outcome: Reproducible (with author involvement)

Substantial implications for the original conclusions: N/A

R Markdown report: https://perma.cc/L86J-CWUT

OSF/Github repository: https://osf.io/ghst4/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.4497763.v1

Description:

The authors provided extensive Matlab code outlining the original
analyses and outputting what appeared to be the primary results
sentences and figures in the paper. However, we found that these did not
actually match up with the numbers reported in the original article. We
noticed that there were some trial exclusions in the main analysis file
that were described as ``optional''. We tried running the code both when
excluding these optional trials and when including them, however,
neither matched the values reported in the paper.

We asked the authors for their input. It turns out that they had already
realised the original code they uploaded did not completely reproduce
the reported findings, and had additionally discovered a mistake in the
data they had posted to the Open Science Framework. They had already
posted corrected data and code to the OSF (https://osf.io/mx92g/) but
our team was not aware of this as it was not mentioned in the paper. The
authors say that they had considered posting a corrigendum to the paper
but decided not to pursue that because Psychological Science only
publishes one if the error ``significantly affects an article's findings
or conclusions or a reader's understanding.'' The authors ``did not feel
this met that threshold''.

A readme file accompanying the updated OSF materials contains further
details about the original errors:

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
The ``Example analysis code'' component uploaded to OSF 10/17/2014
(https://osf.io/7gvh9/) was not sufficiently documented and hard to
use in replicating the published results. The code presented here is a
commented version of the code used to produce the reported analyses,
and therefore identically replications the results from the paper.
\end{enumerate}
\end{quote}

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{1}
\tightlist
\item
The data file with the mouse traces
(sullivanEtAl2014ForPub\_mouse.csv) contained several errors, leading
to another source of discrepancy from the paper's reported statistics.
The mouse tracking data uploaded in this component (newData.csv) is
the original raw data.
\end{enumerate}
\end{quote}

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{2}
\tightlist
\item
This code reports that the percent of trials excluded from analysis is
8.1\%, but the paper reports that it was 13.3\%. This updated 8.1\%
figure is correct, and the reported number in the paper is a mistake."
\end{enumerate}
\end{quote}

With the updated data and code, we are now able to reproduce the values
and have attributed the original reproducibility problems to the errors
in the data file as specified above by the authors.

\hypertarget{vignette-2-article-3-10-2014_ps}{%
\subsection{Vignette 2 (article
3-10-2014\_PS)}\label{vignette-2-article-3-10-2014_ps}}

Outcome: Not fully reproducible (with author involvement)

Substantial implications for the original conclusions: Unknown (analysis
could not be completed)

R Markdown report: https://perma.cc/6FBE-GXE6

OSF/Github repository: https://osf.io/9vqs5/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.7978743.v1

Description:

We were able to reproduce the descriptive results presented in Figure 2
but encountered problems reproducing the inferential statistics. After
receiving a more specific analysis specification from the author, we
were mostly able to reproduce the key 3-way interaction from the
repeated measures ANOVA. Specifically, we were able to reproduce the
degrees of freedom and the p-value. However, we still found a minor
numerical discrepancy in the F-statistic and a major numerical
discrepancy with the \(\eta^2\) (although the absolute magnitude of the
difference was small).

Despite several attempts, we were only able to reproduce a subset of the
follow-up Tukey tests and needed further information to proceed. One
p-value appeared to be a decision error, because it was reported as p
\textgreater{} 0.09 but we obtained p = 0.02.

We requested the following additional information from the authors: (1)
the calculation that they used for the \(\eta^2\) value in the repeated
measures ANOVA; (2) precisely what subsets of the data were used for
each p-value reported in the follow-up Tukey tests. However, we did not
receive a response to this request and have concluded that there is
insufficient information to proceed with the analysis.

\hypertarget{vignette-3-article-1-1-2015_ps}{%
\subsection{Vignette 3 (article
1-1-2015\_PS)}\label{vignette-3-article-1-1-2015_ps}}

Outcome: Not fully reproducible (without author involvement)

Substantial implications for the original conclusions: Unlikely.

R Markdown report: https://perma.cc/MY9V-97XQ

OSF/Github repository: https://osf.io/6p28a/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.7037066.v1

Description:

We could reproduce all target values aside from one major numerical
discrepancy for an F-value. In the results section of the paper it was
stated that all F-values of the non-significant tests were \textless{}1,
but in model 3, the F-value we obtained was 2.21. We contacted the first
author on Apr 25 and both the first author and last author on May 22 to
ask for their input. We've received no response and have therefore
concluded this reproducibility check. Although the cause of
non-reproducibility cannot be identified, this single observed
discrepancy is unlikely to be consequential for the article's original
conclusions.

\hypertarget{vignette-4-article-11-11-2014_ps}{%
\subsection{Vignette 4 (article
11-11-2014\_PS)}\label{vignette-4-article-11-11-2014_ps}}

Outcome: Not fully reproducible (with author involvement)

Substantial implications for the original conclusions: Unlikely

R Markdown report: https://perma.cc/K2Y6-58YK

OSF/Github repository: https://osf.io/huqmg/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.3190579.v1

Description:

This reproducibility check was largely a success. There were some
difficulties mapping the reported analyses to the variable names in the
data file, however these were eventually resolved with a bit of
guesswork.

There was one major numerical discrepancy where a p value was reported
as = .001 but we obtained 3.695211e-13. It seems very likely the
intention was to report p \textless{}.001 as the reported t-value is
consistent with this significance level (the authors have confirmed over
email that it was probably a typo). This single observed discrepancy is
unlikely to be consequential for the article's original conclusions.

\hypertarget{vignette-5-article-11-12-2014_ps}{%
\subsection{Vignette 5 (article
11-12-2014\_PS)}\label{vignette-5-article-11-12-2014_ps}}

Outcome: Reproducible (without author involvement)

Substantial implications for the original conclusions: N/A

R Markdown report: https://perma.cc/29RS-Y55L

OSF/Github repository: https://osf.io/btpnm/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.5508100.v1

Description:

We encountered no major difficulties reproducing all target values for
this article.

\hypertarget{vignette-6-article-16-11-2014_ps}{%
\subsection{Vignette 6 (article
16-11-2014\_PS)}\label{vignette-6-article-16-11-2014_ps}}

Outcome: Reproducible (without author involvement)

Substantial implications for the original conclusions: N/A

R Markdown report: https://perma.cc/USK3-244C

OSF/Github repository: https://osf.io/ubwjm/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.2068771.v1

Description:

The inferential statistical tests employed were not explicitly named in
the original article. However, we made an educated guess that they were
chi-squared tests given the context, and this enabled us to reproduce
all target values.

\hypertarget{vignette-7-article-1-6-2014_ps}{%
\subsection{Vignette 7 (article
1-6-2014\_PS)}\label{vignette-7-article-1-6-2014_ps}}

Outcome: Not fully reproducible (with author involvement)

Substantial implications for the original conclusions: Unlikely.

R Markdown report: https://perma.cc/89A9-QLXN

OSF/Github repository: https://osf.io/4k6v5/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.5582627.v1

Description:

Note that for this reproducibility check, a corrigendum had previously
been published for the target article, so we aimed to reproduce the
values reported in the corrigendum rather than the original article. We
also could not work out how to implement the inferential analyses in R
and used SPSS instead - aided by the SPSS syntax provided by the
original authors.

All inferential statistics could be reproduced using SPSS. However, we
found one major numerical discrepancy in the descriptive statistics
(which were run in R): a reported mean of 0.05, which according to our
analysis was 0.005. The relevant value also seemed to be 0.005 according
to Figure 1 of the corrigendum. We contacted one of the authors and they
confirmed 0.005 is the correct value and the value reported in the
corrigendum is a typo. This single observed discrepancy is unlikely to
be consequential for the article's original conclusions.

\hypertarget{vignette-8-article-10-7-2014_ps}{%
\subsection{Vignette 8 (article
10-7-2014\_PS)}\label{vignette-8-article-10-7-2014_ps}}

Outcome: Reproducible (with author involvement)

Substantial implications for the original conclusions: N/A

R Markdown report: https://perma.cc/53BJ-BP3P

OSF/Github repository: https://osf.io/5eps7/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.4524649.v1

Description:

We could initially reproduce all values except for one p-value reported
for a simple effects analysis. After the original authors shared SPSS
syntax with us it became clear that they were implementing the analysis
in a different way to what we had understood from reading the paper.
When we used the original method, we were able to reproduce the p-value
successfully.

\hypertarget{vignette-9-article-16-2-2015_ps}{%
\subsection{Vignette 9 (article
16-2-2015\_PS)}\label{vignette-9-article-16-2-2015_ps}}

Outcome: Not fully reproducible (without author involvement)

Substantial implications for the original conclusions: Unlikely.

R Markdown report: https://perma.cc/BD6J-U6V4

OSF/Github repository: https://osf.io/r9j83/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.2045757.v1

Description:

We could successfully reproduce all reported descriptive statistics for
the target values; however, we observed several major numerical
discrepancies for correlation coefficients and their confidence
intervals. However, although these cases exceeded our 10\% percentage
error threshold, the absolute magnitude of the difference was small (all
\textless{} .03 correlation units). It is not clear to us what the
source of the discrepancy is, although it could be that different
software packages (the authors used SAS and we are using R) give
different values for these tests.

We attempted to run the SAS program which was provided alongside the
article. However, the program is trying to read in the following text
files: `explicit.txt' `sessions.txt', `sessionTasks.txt', `iat.txt',
which, as far as we can tell, are not shared along with the article. We
contacted the corresponding author on April 23, 2019, to request these
files. We initially received a response from the corresponding author on
Apr 23 (2019) which offered to help in principle but noted that this
could take some time. We sent follow-up emails on April 25, May 22, and
July 2, but received no further response. We ultimately decided to close
the reproducibility check. Note that although we made contact with an
author, no substantive assistance was provided - this is reflected in
the categorisation of this case as ``not fully reproducible (no author
involvement)''

Although we cannot identify the causal locus of non-reproducibility, the
magnitude of the differences suggests that none of the issues we
identified are likely to substantially change the authors' original
conclusions.

\hypertarget{vignette-10-article-16-9-2014_ps}{%
\subsection{Vignette 10 (article
16-9-2014\_PS)}\label{vignette-10-article-16-9-2014_ps}}

Outcome: Reproducible (without author involvement)

Substantial implications for the original conclusions: N/A

R Markdown report: https://perma.cc/K5LB-WD45s

OSF/Github repository: https://osf.io/hz8mn/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.1533019.v1

Description:

We encountered some numerical differences for sample sizes reported in
the original article compared to our reanalysis, however these
differences did not meet the threshold to be classified as major
numerical discrepancies. We could successfully reproduce all target
values.

\hypertarget{vignette-11-article-3-4-2015_ps}{%
\subsection{Vignette 11 (article
3-4-2015\_PS)}\label{vignette-11-article-3-4-2015_ps}}

Outcome: Not fully reproducible (with author involvement)

Substantial implications for the original conclusions:

R Markdown report: https://perma.cc/PGT4-GN8U

OSF/Github repository: https://osf.io/zkmgw/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.6468572.v1

Description:

In this reproducibility check we were able to reproduce the descriptive
statistics (Figure 3) but ran into difficulties reproducing some of the
inferential statistics. We contacted the authors and received assistance
that resolved some issues. Specifically, information was provided about
the identity of statistical tests, units of analysis, and normalization
procedures, that were not stated in the paper. Unfortunately, we then
lost contact with the authors and some reproducibility issues remained
that we could not resolve.

We could not reproduce two effect sizes (d). The authors reported that
they tried to reproduce these values themselves and could ``closely
reproduce the Cohen's d values''. However, they did not share the exact
values they obtained in their re-analysis. We could also not reproduce
one mean and one bound of a confidence interval. The causal locus of
these issues is unclear.

The reproducibility issues do not appear to undermine the original
conclusions. The obtained effect sizes are smaller than those reported,
but not substantially so. The mean and ci discrepancies are of low
magnitude.

\hypertarget{vignette-12-article-2-10-2014_ps}{%
\subsection{Vignette 12 (article
2-10-2014\_PS)}\label{vignette-12-article-2-10-2014_ps}}

Outcome: Reproducible (without author involvement)

Substantial implications for the original conclusions: N/A

R Markdown report: https://perma.cc/V4R9-V8JJ

OSF/Github repository: https://osf.io/2mq46/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.7548371.v1

Description:

The manuscript did not specify the way that confidence intervals were
computed and it was necessary to make an educated guess. All target
values were reproduced successfully.

\hypertarget{vignette-13-article-3-9-2014_ps}{%
\subsection{Vignette 13 (article
3-9-2014\_PS)}\label{vignette-13-article-3-9-2014_ps}}

Outcome: Not fully reproducible (with author involvement)

Substantial implications for the original conclusions: Unlikely

R Markdown report: https://perma.cc/4PFB-LHBP

OSF/Github repository: https://osf.io/b3v7h/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.4160541.v1

Description:

Whilst trying to reproduce an ANOVA we noted that two participants
(participant 25 for both the Moroccan and Spaniard samples) had missing
data for the PAST condition. However, the degrees of freedom reported in
the paper appeared to reflect that the participants \emph{were} included
in this analysis. We asked the original authors for clarification and
they informed us that there were typos in the data file they had posted
on OSF. The wrong participant ID had been added for some lines of data,
making it appear that two participants were missing data. Once the typos
were corrected, the degrees of freedom in our re-analyses matched those
reported in the paper.

Two major numerical discrepancies with p-values remained. In both cases,
the authors reported that p = .001, but the values we obtained were
lower than this. We discussed this issue with the authors and they said
that it is their policy to report p values that are \textless{}.001 as p
= .001. Unfortunately this approach guarantees inaccurate reporting.
According to the APA, researchers should ``report p values less than
.001 as p \textless{} .001'' (APA, 2009; p.~114). The reproducibility
issues we encountered do not appear to undermine the conclusions drawn
in the original article.

\hypertarget{vignette-14-article-4-1-2015_ps}{%
\subsection{Vignette 14 (article
4-1-2015\_PS)}\label{vignette-14-article-4-1-2015_ps}}

Outcome: Reproducible (with author involvement)

Substantial implications for the original conclusions: N/A

R Markdown report: https://perma.cc/R2V2-Y3HS

OSF/Github repository: https://osf.io/c6u95/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.4555168.v1

Description:

We initially could not reproduce some degrees of freedom and F-values
despite trying multiple different model specifications. We contacted the
authors for assistance and they correctly pointed out that we had missed
an important footnote (footnote 2) which read: ``The sphericity
assumption for this analysis and the corresponding analysis in
Experiments 2 and 3 was not met. We report multivariate test results, as
recommended by Maxwell and Delaney (2004), because these tests are more
optimal than correcting for sphericity. The pattern of results did not
differ depending on whether we used either a multivariate test or
sphericity correction.'' They also sent SPSS syntax and a screenshot of
the output - which demonstrated successful reproduction of the target
values.

Unfortunately at this time our team does not have access to SPSS nor the
expertise to independently implement these multivariate tests in R (the
analysis appears to be beyond our operational definition of a
`reasonably straightforward analysis'), however the SPSS syntax and
output indicates that these outcomes are reproducible.

\hypertarget{vignette-15-article-4-11-2014_ps}{%
\subsection{Vignette 15 (article
4-11-2014\_PS)}\label{vignette-15-article-4-11-2014_ps}}

Outcome: Reproducible (without author involvement)

Substantial implications for the original conclusions: N/A

R Markdown report: https://perma.cc/VA79-D42D

OSF/Github repository: https://osf.io/2kz9b/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.2298575.v1

Description:

We encountered no major difficulties reproducing all target values for
this article.

\hypertarget{vignette-16-article-5-4-2015_ps}{%
\subsection{Vignette 16 (article
5-4-2015\_PS)}\label{vignette-16-article-5-4-2015_ps}}

Outcome: Reproducible (with author involvement)

Substantial implications for the original conclusions: N/A

R Markdown report: https://perma.cc/9DEH-7YPY

OSF/Github repository: https://osf.io/bm4cg/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.1310152.v1

Description:

Although it was specified in the paper that one participant was excluded
from the analyses due to a computer malfunction, the data files provided
online did not specify which participant this referred to. Thus, prior
to communication with the author, we were not able to proceed with the
analyses. After the author identified the participant to be excluded, we
were able to reproduce all target values.

\hypertarget{vignette-17-article-6-1-2015_ps}{%
\subsection{Vignette 17 (article
6-1-2015\_PS)}\label{vignette-17-article-6-1-2015_ps}}

Outcome: Reproducible (without author involvement)

Substantial implications for the original conclusions: N/A

R Markdown report: https://perma.cc/BTS6-5MH6

OSF/Github repository: https://osf.io/f9q28/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.0405521.v1

Description:

We encountered no major difficulties reproducing all target values for
this article.

\hypertarget{vignette-18-article-6-7-2014_ps}{%
\subsection{Vignette 18 (article
6-7-2014\_PS)}\label{vignette-18-article-6-7-2014_ps}}

Outcome: Reproducible (without author involvement)

Substantial implications for the original conclusions: N/A

R Markdown report: https://perma.cc/Z8SP-CEDV

OSF/Github repository: https://osf.io/qezax/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.0223565.v1

Description:

We encountered no major difficulties reproducing all target values for
this article.

\hypertarget{vignette-19-article-7-3-2015_ps}{%
\subsection{Vignette 19 (article
7-3-2015\_PS)}\label{vignette-19-article-7-3-2015_ps}}

Outcome: Not fully reproducible (without author involvement)

Substantial implications for the original conclusions: Unknown (analysis
could not be completed)

R Markdown report: https://perma.cc/77ND-XHKN

OSF/Github repository: https://osf.io/jgp8e/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.7977267.v1

Description:

In our initial attempts, we found a number of minor numerical
discrepancies in the descriptive statistics whether we used medians or
means. We then encountered major numerical discrepancies in the
inferential statistics. There were a number of aspects of the original
analysis and data files we were unclear about. Unfortunately despite
emailing the original authors several times we have not received a
response to our questions (we have received responses saying they will
get back to us, but this has not happened \textgreater{} 20 months after
the last message). We have thus classified these issues as `insufficient
information errors'. The issues we are unclear about are outlined in
full detail in the reproducibility report. In brief, they refer to
unclear labels and codings for variables in the data file, unclear use
of means vs.~medians, unclear levels of aggregation. Despite trying out
multiple combinations of these factors based on educated guesses, we
remain unable to reproduce some values and cannot complete parts of the
analysis and it is unclear why.

\hypertarget{vignette-20-article-8-12-2014_ps}{%
\subsection{Vignette 20 (article
8-12-2014\_PS)}\label{vignette-20-article-8-12-2014_ps}}

Outcome: Not fully reproducible (with author involvement)

Substantial implications for the original conclusions: Unlikely

R Markdown report: https://perma.cc/R38G-QQKG

OSF/Github repository: https://osf.io/5hv6f/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.0413564.v1

Description:

We initially ran into a problem reproducing the exclusions reported in
the paper. We used both the SPSS and Matlab files provided by the
authors and attempted to apply the exclusions in several different ways.
As all were unsuccessful, we contacted the original authors to ask for
assistance. They created a video of themselves performing the analysis
in SPSS and reproducing the exclusion numbers reported in the paper.
Watching this video helped us to identify that we had misunderstood the
sentence in the paper that reported the exclusions, specifically: ``Only
participants for whom both d'1 and d'2 could be computed for the
analysis subset (i.e., who had non-zero counts in every cell) were
included; above chance: n = 165, at chance: n = 33.'' We had excluded
participants with missing values in any of the cells, whereas the
original analysis excluded participants with missing values in cells
relevant to the analysis subset. After being sent the video, we were
able to reproduce this part of the analysis successfully and proceeded
with the remainder of the analysis.

We were unable to reproduce two values - a t-value (0.17 reported
vs.~0.12 in our analysis) and a d-value (0.03 reported vs 0.02 in our
analysis). We contacted the authors again and they said they also could
not reproduce these values either - in their own re-analyses they obtain
the same values as we do. The reason for the differences is unknown. The
authors suggest it might be due to a change in version of the SPSS
software that was used to run the original analysis. Although they
qualify as `major numerical discrepancies according to our operational
definition, the magnitude of these differences indicates that they are
unlikely to be consequential to the authors' original conclusions.

\hypertarget{vignette-21-article-8-5-2015_ps}{%
\subsection{Vignette 21 (article
8-5-2015\_PS)}\label{vignette-21-article-8-5-2015_ps}}

Outcome: Reproducible (without author involvement)

Substantial implications for the original conclusions: N/A

R Markdown report: https://perma.cc/WN37-9T2K

OSF/Github repository: https://osf.io/uaeqw/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.2646134.v1

Description:

We encountered no major difficulties reproducing all target values for
this article.

\hypertarget{vignette-22-article-8-7-2014_ps}{%
\subsection{Vignette 22 (article
8-7-2014\_PS)}\label{vignette-22-article-8-7-2014_ps}}

Outcome: Reproducible (with author involvement)

Substantial implications for the original conclusions: N/A

R Markdown report: https://perma.cc/4779-EB7S

OSF/Github repository: https://osf.io/mxwq7/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.7903145.v1

Description:

Initially we encountered major numerical discrepancies for the analysis
of gain scores. The t-value and the effect size obtained were
considerably smaller compared to the ones reported in the paper. We
originally calculated the gain scores by averaging scores in each test
phase (pre and post) in each condition for each subject and then
subtracting the averages for the post test from the ones for the pre
test. This way of calculating scores was based on the paper's
description: ``Gain scores were calculated by subtracting each
participant's pretest score from his or her posttest score.''

We contacted the original authors and they informed us that the gain
scores were calculated in a different way, namely by subtracting pre
from post test scores for each specific lesson (e.g., plate tectonics)
and then averaging across lesson difference scores to get a mean gain
score per subject and condition. This also included excluding some
lessons for some of the participants due to absence. This way of
calculating the gain scores was not obvious based on the information in
the paper. The supplementary material also did not include any
additional information about gain scores. After correcting the way the
gain scores were calculated we were able to reproduce all target values
successfully.

\hypertarget{vignette-23-article-8-8-2014_ps}{%
\subsection{Vignette 23 (article
8-8-2014\_PS)}\label{vignette-23-article-8-8-2014_ps}}

Outcome: Reproducible (with author involvement)

Substantial implications for the original conclusions: N/A

R Markdown report: https://perma.cc/58C5-HWDQ

OSF/Github repository: https://osf.io/buegj/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.2241042.v1

Description:

We initially had trouble reproducing the principle components analysis
to generate the wise reasoning composite scores from the raw data due to
a missing variable for question 4A from Table 1 and 5 missing values in
the raw data file that were not mentioned in the paper. But after author
clarifications, these issues were resolved.

Some major numerical discrepancies remained, though their absolute
magnitude was low. The authors provided SPSS syntax and we were able to
implement this and reproduce the values reported in the paper. Despite
extensive efforts we were unable to localize the exact reason why we
could not reproduce the values in R. Our suspicion is that these might
be caused by opaque SPSS/R computational differences. Nevertheless, we
were able to reproduce all target values using a combination of R and
SPSS.

\hypertarget{vignette-24-article-9-2-2015_ps}{%
\subsection{Vignette 24 (article
9-2-2015\_PS)}\label{vignette-24-article-9-2-2015_ps}}

Outcome: Reproducible (without author involvement)

Substantial implications for the original conclusions: N/A

R Markdown report: https://perma.cc/9VCW-MXVX

OSF/Github repository: https://osf.io/cneqr/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.7685589.v1

Description:

We encountered no major difficulties reproducing all target values for
this article.

\hypertarget{vignette-25-article-9-5-2014_ps}{%
\subsection{Vignette 25 (article
9-5-2014\_PS)}\label{vignette-25-article-9-5-2014_ps}}

Outcome: Not fully reproducible (with author involvement)

Substantial implications for the original conclusions: Unknown (analysis
could not be completed)

R Markdown report: https://perma.cc/SVH6-8EN6

OSF/Github repository: https://osf.io/fam6g/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.8022934.v1

Description:

We were initially able to reproduce the descriptive statistics and the
three key figures. But there were four insufficient information errors
pertaining to the original analysis specifications that prevented a full
reproducibility check (see reproducibility report for details). We
contacted the original authors for assistance and they clarified some
aspects of the analyses which resolved some issues.

However, we continued to have difficulties with the section reporting
post-hoc comparisons. There were three major numerical discrepancies for
p-values. The authors report using `Sidak corrections' but it is unclear
if they were applied to the alpha threshold (correctly) or the p-values
themselves. If corrections were applied to the p values, then this could
potentially explain the differences, but the article does not identify
the family of hypotheses for which corrections were applied. So we also
cannot correct the alpha threshold. We have decided to report the major
numerical discrepancies for the p values, but also record an
insufficient information error as we cannot determine whether these were
decision errors or not. Additionally, for one degree of freedom, we
obtained 4 whereas the article reports 3. This may be a typo but the
cause is not clear.

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-kidwell2016}{}%
Kidwell, M. C., Lazarevi\a'c, L. B., Baranski, E., Hardwicke, T. E.,
Piechowski, S., Falkenberg, L.-S., Kennett, C., Slowik, A., Sonnleitner,
C., Hess-Holden, C., Errington, T. M., Fiedler, S., \& Nosek, B. A.
(2016). Badges to Acknowledge Open Practices: A Simple, Low-Cost,
Effective Method for Increasing Transparency. \emph{PLOS Biology},
\emph{14}(5), e1002456.
\url{https://doi.org/10.1371/journal.pbio.1002456}
\end{appendix}

\end{document}