### 2020-06-08 manuscript/supplementary_information.tex

\clearpage
\makeatletter
\efloat@restorefloats
\makeatother

\begin{appendix}
\hypertarget{sup_sample}{%
\section{Supplementary sample information}\label{sup_sample}}

We downloaded data from the Kidwell et al. (2016) study (available here:
https://osf.io/u6g7t/) and selected the 47 articles that had received an
`Open Data Badge'. Kidwell and colleagues had determined that 35 of
these articles were accompanied by data sets that were accessible,
correct, complete, and understandable. One team member (TEH) examined
each of these articles and attempted to identify a `substantive finding'
supported by a `relatively straightforward analysis' according to the
following operational definitions (adopted directly from Hardwicke et
al.~2018): (A) A `substantive finding' is one that supports a central
conclusion of the article under scrutiny. This is to some extent
subjective, but generally refers to findings supported in the article's
abstract and/or displayed in a figure/table; (B) `Relatively
straightforward analyses' involved behavioural data only and employed
quantitative techniques that would be commonly found in an introductory
level statistics textbook aimed at undergraduate psychology students
(e.g., {\textbf{???}}). Examples included means, medians, standard
deviations, confidence intervals, standardized effect sizes,
correlations, t-tests, and ANOVAs.

The first finding encountered in the article that met these eligibility
criteria was selected. The target values for the reproducibility checks
were a coherent set of descriptive and inferential statistics related to
this finding, typically 2-3 paragraphs of information, occasionally
accompanied by a table or figure. During this assessment, 9 articles
were excluded because they did not meet the eligibility criteria (ID
codes: 12-8-2014 PS, 21-7-2014 PS, 13-6-2014 PS, 18-7-2014 PS, 3-2-2015
PS, 6-3-2015 PS, 16-8-2014 PS, 19-4-2015 PS, 15-2-2015 PS). An
additional article was excluded because the data no longer appeared to
be available (contra Kidwell et al., 2016; ID code: 8-11-2014 PS). In
total, 25 eligible articles were identified. A bibliography file is
available where article ID codes can be matched to their references
(https://osf.io/rszey/). Note that there was a reporting error in our
pre-registered protocol which said that article 16-2-2015 was excluded
when it was not.

A precision analysis (not pre-registered and performed after study
completion) shows the range of expected margin of error (confidence
interval width) for the range of possible effect sizes given our sample
size of 25 articles (Figure @ref(fig:precisionCurve)). The maximum
margin of error is 0.20.

\begin{figure}
\centering
\includegraphics{figs/precisionCurve-1.pdf}
\caption{(\#fig:precisionCurve)Precision analysis yielding expected
margin of error (confidence interval width) relative to effect size
(proportion) for a sample size of 25.}
\end{figure}

\hypertarget{sup_design}{%
\section{Supplementary design and procedures}\label{sup_design}}

The most serious consequence of non-reproducibility is that it
undermines the credibility of associated scientific claims. We therefore
attempted to determine the extent to which any reproducibility issues we
encountered had substantial implications for related conclusions drawn
in the original papers. As noted in a previous investigation (Hardwicke
et al., 2018), this is a complex issue that cannot be straightforwardly
measured by a quantitative index. We therefore, considered this question
on a case by case basis, using multiple factors to inform our judgement,
including the presence/absence of decision errors, the quantity of and
type of target values that could not be reproduced, the difference in
magnitude of key values like effect sizes, and the specificity of the
hypothesis under scrutiny. This was a necessarily subjective exercise;
however the judgement for each reproducibility check was agreed upon by
the (two or more) team members who worked on it in addition to the first
author (TEH) and senior author (MCF). The rationale underlying each
judgement is outlined in the reproducibility vignettes in Supplementary
Materials @ref(vignettes).

Determining the causal locus of non-reproducibility was not
straightforward because we often did not have access to comprehensive
information about the original analysis process. Additionally, there is
not necessarily a direct mapping between discrete causal loci and
reproducibility issues; multiple numerical discrepancies could be
attributable to a single cause or multiple causes. Sometimes the causal
locus of non-reproducibility became apparent after discussions with the
original authors, however, all classifications were made independently
by our team.

Comparisons between values were partly automated through the use of a
custom R function (https://osf.io/teuzb/) that accounted for rounding
errors by rounding new values to the same number of decimal places as
old values prior to computing the percentage error and classifying the
value as a match, minor discrepancy, or major discrepancy. Note that for
212 values reported relative to a threshold (e.g., p \textless{} 0.05)
or embedded in figures, we could not directly calculate percentage error
and it was necessary to `eyeball' (i.e., visually compare side-by-side)
the values to confirm a match.

All team members had started or completed graduate level training in
psychology and had experience conducting the types of data analyses
typically found in an introductory psychology statistics textbook (e.g.,
Field et al., 2012).

In the first contact emails we sent to original authors, we explained
the project, the issues we had encountered with their article, and
provided a copy of an interim reproducibility report. If we were
notified that the email was not delivered, we spent up to 5 minutes
searching for a more recent contact email address. If the corresponding
author(s) did not respond after two weeks, a reminder email was sent.

\hypertarget{sup_results}{%
\section{Supplementary results}\label{sup_results}}

Ninety-five per cent confidence intervals (CIs) displayed in square
brackets are based on the Wilson method with continuity correction for
binomial proportions ({\textbf{???}}) and the Sison--Glaz method for
multinomial proportions ({\textbf{???}}).

\begin{figure}
\centering
\includegraphics{figs/locusPlot-1.pdf}
\caption{(\#fig:locusPlot)Frequency of discrete causal loci underlying
non-reproducibility across all reproducibility checks. Note that some
articles contained multiple causal loci.}
\end{figure}

\begin{figure}
\centering
\includegraphics{figs/timePlot-1.pdf}
\caption{(\#fig:timePlot)Estimated number of active work hours spent by
our team on each reproducibility check, presented separately by
reproducibility outcome.}
\end{figure}

Six articles were accompanied by analysis code, including one with
Matlab scripts and SPSS syntax (8-12-2014\_PS), one with Matlab scripts
only (2-2-2015\_PS), two with SPSS syntax only (2-10-2014\_PS,
16-9-2014\_PS), one with an R script (8-5-2015\_PS), and one with SAS
syntax (16-2-2015\_PS). Three articles were reproducible without author
involvement; one article was reproducible with author involvement; one
article was not reproducible with author involvement; and one article
was not reproducible with no author response. Team members estimated
that they spent between 3 and 10 (median = 6.50, interquartile range =
3) hours actively working on each of the reproducibility checks for
articles with scripts.

\hypertarget{prereg}{%
\section{Deviations from pre-registered protocol}\label{prereg}}

The pre-registered protocol (https://osf.io/2cnkq/) does not state that
we will report confidence intervals. However, we decided to report
confidence intervals as this would aid inferences beyond the sample.

The pre-registered protocol used the terminology ``major/minor errors''.
To improve clarity, we use ``major/minor numerical discrepancies''.

The pre-registered protocol used the terminology ``reproducibility
failure'' and ``reproducibility success'' in reference to article
classifications. To improve clarity, we use ``not fully reproducible''
and ``reproducible''.

The pre-registered protocol stated that ``After contacting authors for
assistance, a maximum time-limit of 2 months for resolution of any
issues will be observed. Any reproducibility issues that cannot be
resolved within this time-period will be considered reproducibility
failures for the purposes of the present investigation.'' We decided not
to impose this time limit as it seemed unreasonable that delays on our
side could influence reproducibility outcomes. Authors were able to
respond until the project was completed on June 2, 2020.

\hypertarget{vignettes}{%
\section{Reproducibility vignettes}\label{vignettes}}

Each of the reproducibility checks is summarised below in a short
vignette. Each vignette contains a link to a full R Markdown report
rendered in HTML that will take the reader through the reanalysis
process step-by-step. A link is also provided to an OSF/Github
repository that contains all relevant data and analysis code and a link
to a Code Ocean container which recreates the software environment in
which the original analyses were run to facilitate reproducibility.

\hypertarget{vignette-1-article-2-2-2015_ps}{%
\subsection{Vignette 1 (article
2-2-2015\_PS)}\label{vignette-1-article-2-2-2015_ps}}

Outcome: Reproducible (with author involvement)

Substantial implications for the original conclusions: N/A

R Markdown report: https://perma.cc/L86J-CWUT

OSF/Github repository: https://osf.io/ghst4/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.4497763.v1

Description:

The authors provided extensive Matlab code outlining the original
analyses and outputting what appeared to be the primary results
sentences and figures in the paper. However, we found that these did not
actually match up with the numbers reported in the original article. We
noticed that there were some trial exclusions in the main analysis file
that were described as ``optional''. We tried running the code both when
excluding these optional trials and when including them, however,
neither matched the values reported in the paper.

We asked the authors for their input. It turns out that they had already
realised the original code they uploaded did not completely reproduce
the reported findings, and had additionally discovered a mistake in the
data they had posted to the Open Science Framework. They had already
posted corrected data and code to the OSF (https://osf.io/mx92g/) but
our team was not aware of this as it was not mentioned in the paper. The
authors say that they had considered posting a corrigendum to the paper
but decided not to pursue that because Psychological Science only
publishes one if the error ``significantly affects an article's findings
or conclusions or a reader's understanding.'' The authors ``did not feel
this met that threshold''.

A readme file accompanying the updated OSF materials contains further
details about the original errors:

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  The ``Example analysis code'' component uploaded to OSF 10/17/2014
  (https://osf.io/7gvh9/) was not sufficiently documented and hard to
  use in replicating the published results. The code presented here is a
  commented version of the code used to produce the reported analyses,
  and therefore identically replications the results from the paper.
\end{enumerate}
\end{quote}

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{1}
\tightlist
\item
  The data file with the mouse traces
  (sullivanEtAl2014ForPub\_mouse.csv) contained several errors, leading
  to another source of discrepancy from the paper's reported statistics.
  The mouse tracking data uploaded in this component (newData.csv) is
  the original raw data.
\end{enumerate}
\end{quote}

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{2}
\tightlist
\item
  This code reports that the percent of trials excluded from analysis is
  8.1\%, but the paper reports that it was 13.3\%. This updated 8.1\%
  figure is correct, and the reported number in the paper is a mistake."
\end{enumerate}
\end{quote}

With the updated data and code, we are now able to reproduce the values
and have attributed the original reproducibility problems to the errors
in the data file as specified above by the authors.

\hypertarget{vignette-2-article-3-10-2014_ps}{%
\subsection{Vignette 2 (article
3-10-2014\_PS)}\label{vignette-2-article-3-10-2014_ps}}

Outcome: Not fully reproducible (with author involvement)

Substantial implications for the original conclusions: Unknown (analysis
could not be completed)

R Markdown report: https://perma.cc/6FBE-GXE6

OSF/Github repository: https://osf.io/9vqs5/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.7978743.v1

Description:

We were able to reproduce the descriptive results presented in Figure 2
but encountered problems reproducing the inferential statistics. After
receiving a more specific analysis specification from the author, we
were mostly able to reproduce the key 3-way interaction from the
repeated measures ANOVA. Specifically, we were able to reproduce the
degrees of freedom and the p-value. However, we still found a minor
numerical discrepancy in the F-statistic and a major numerical
discrepancy with the \(\eta^2\) (although the absolute magnitude of the
difference was small).

Despite several attempts, we were only able to reproduce a subset of the
follow-up Tukey tests and needed further information to proceed. One
p-value appeared to be a decision error, because it was reported as p
\textgreater{} 0.09 but we obtained p = 0.02.

We requested the following additional information from the authors: (1)
the calculation that they used for the \(\eta^2\) value in the repeated
measures ANOVA; (2) precisely what subsets of the data were used for
each p-value reported in the follow-up Tukey tests. However, we did not
receive a response to this request and have concluded that there is
insufficient information to proceed with the analysis.

\hypertarget{vignette-3-article-1-1-2015_ps}{%
\subsection{Vignette 3 (article
1-1-2015\_PS)}\label{vignette-3-article-1-1-2015_ps}}

Outcome: Not fully reproducible (without author involvement)

Substantial implications for the original conclusions: Unlikely.

R Markdown report: https://perma.cc/MY9V-97XQ

OSF/Github repository: https://osf.io/6p28a/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.7037066.v1

Description:

We could reproduce all target values aside from one major numerical
discrepancy for an F-value. In the results section of the paper it was
stated that all F-values of the non-significant tests were \textless{}1,
but in model 3, the F-value we obtained was 2.21. We contacted the first
author on Apr 25 and both the first author and last author on May 22 to
ask for their input. We've received no response and have therefore
concluded this reproducibility check. Although the cause of
non-reproducibility cannot be identified, this single observed
discrepancy is unlikely to be consequential for the article's original
conclusions.

\hypertarget{vignette-4-article-11-11-2014_ps}{%
\subsection{Vignette 4 (article
11-11-2014\_PS)}\label{vignette-4-article-11-11-2014_ps}}

Outcome: Not fully reproducible (with author involvement)

Substantial implications for the original conclusions: Unlikely

R Markdown report: https://perma.cc/K2Y6-58YK

OSF/Github repository: https://osf.io/huqmg/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.3190579.v1

Description:

This reproducibility check was largely a success. There were some
difficulties mapping the reported analyses to the variable names in the
data file, however these were eventually resolved with a bit of
guesswork.

There was one major numerical discrepancy where a p value was reported
as = .001 but we obtained 3.695211e-13. It seems very likely the
intention was to report p \textless{}.001 as the reported t-value is
consistent with this significance level (the authors have confirmed over
email that it was probably a typo). This single observed discrepancy is
unlikely to be consequential for the article's original conclusions.

\hypertarget{vignette-5-article-11-12-2014_ps}{%
\subsection{Vignette 5 (article
11-12-2014\_PS)}\label{vignette-5-article-11-12-2014_ps}}

Outcome: Reproducible (without author involvement)

Substantial implications for the original conclusions: N/A

R Markdown report: https://perma.cc/29RS-Y55L

OSF/Github repository: https://osf.io/btpnm/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.5508100.v1

Description:

We encountered no major difficulties reproducing all target values for
this article.

\hypertarget{vignette-6-article-16-11-2014_ps}{%
\subsection{Vignette 6 (article
16-11-2014\_PS)}\label{vignette-6-article-16-11-2014_ps}}

Outcome: Reproducible (without author involvement)

Substantial implications for the original conclusions: N/A

R Markdown report: https://perma.cc/USK3-244C

OSF/Github repository: https://osf.io/ubwjm/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.2068771.v1

Description:

The inferential statistical tests employed were not explicitly named in
the original article. However, we made an educated guess that they were
chi-squared tests given the context, and this enabled us to reproduce
all target values.

\hypertarget{vignette-7-article-1-6-2014_ps}{%
\subsection{Vignette 7 (article
1-6-2014\_PS)}\label{vignette-7-article-1-6-2014_ps}}

Outcome: Not fully reproducible (with author involvement)

Substantial implications for the original conclusions: Unlikely.

R Markdown report: https://perma.cc/89A9-QLXN

OSF/Github repository: https://osf.io/4k6v5/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.5582627.v1

Description:

Note that for this reproducibility check, a corrigendum had previously
been published for the target article, so we aimed to reproduce the
values reported in the corrigendum rather than the original article. We
also could not work out how to implement the inferential analyses in R
and used SPSS instead - aided by the SPSS syntax provided by the
original authors.

All inferential statistics could be reproduced using SPSS. However, we
found one major numerical discrepancy in the descriptive statistics
(which were run in R): a reported mean of 0.05, which according to our
analysis was 0.005. The relevant value also seemed to be 0.005 according
to Figure 1 of the corrigendum. We contacted one of the authors and they
confirmed 0.005 is the correct value and the value reported in the
corrigendum is a typo. This single observed discrepancy is unlikely to
be consequential for the article's original conclusions.

\hypertarget{vignette-8-article-10-7-2014_ps}{%
\subsection{Vignette 8 (article
10-7-2014\_PS)}\label{vignette-8-article-10-7-2014_ps}}

Outcome: Reproducible (with author involvement)

Substantial implications for the original conclusions: N/A

R Markdown report: https://perma.cc/53BJ-BP3P

OSF/Github repository: https://osf.io/5eps7/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.4524649.v1

Description:

We could initially reproduce all values except for one p-value reported
for a simple effects analysis. After the original authors shared SPSS
syntax with us it became clear that they were implementing the analysis
in a different way to what we had understood from reading the paper.
When we used the original method, we were able to reproduce the p-value
successfully.

\hypertarget{vignette-9-article-16-2-2015_ps}{%
\subsection{Vignette 9 (article
16-2-2015\_PS)}\label{vignette-9-article-16-2-2015_ps}}

Outcome: Not fully reproducible (without author involvement)

Substantial implications for the original conclusions: Unlikely.

R Markdown report: https://perma.cc/BD6J-U6V4

OSF/Github repository: https://osf.io/r9j83/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.2045757.v1

Description:

We could successfully reproduce all reported descriptive statistics for
the target values; however, we observed several major numerical
discrepancies for correlation coefficients and their confidence
intervals. However, although these cases exceeded our 10\% percentage
error threshold, the absolute magnitude of the difference was small (all
\textless{} .03 correlation units). It is not clear to us what the
source of the discrepancy is, although it could be that different
software packages (the authors used SAS and we are using R) give
different values for these tests.

We attempted to run the SAS program which was provided alongside the
article. However, the program is trying to read in the following text
files: `explicit.txt' `sessions.txt', `sessionTasks.txt', `iat.txt',
which, as far as we can tell, are not shared along with the article. We
contacted the corresponding author on April 23, 2019, to request these
files. We initially received a response from the corresponding author on
Apr 23 (2019) which offered to help in principle but noted that this
could take some time. We sent follow-up emails on April 25, May 22, and
July 2, but received no further response. We ultimately decided to close
the reproducibility check. Note that although we made contact with an
author, no substantive assistance was provided - this is reflected in
the categorisation of this case as ``not fully reproducible (no author
involvement)''

Although we cannot identify the causal locus of non-reproducibility, the
magnitude of the differences suggests that none of the issues we
identified are likely to substantially change the authors' original
conclusions.

\hypertarget{vignette-10-article-16-9-2014_ps}{%
\subsection{Vignette 10 (article
16-9-2014\_PS)}\label{vignette-10-article-16-9-2014_ps}}

Outcome: Reproducible (without author involvement)

Substantial implications for the original conclusions: N/A

R Markdown report: https://perma.cc/K5LB-WD45s

OSF/Github repository: https://osf.io/hz8mn/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.1533019.v1

Description:

We encountered some numerical differences for sample sizes reported in
the original article compared to our reanalysis, however these
differences did not meet the threshold to be classified as major
numerical discrepancies. We could successfully reproduce all target
values.

\hypertarget{vignette-11-article-3-4-2015_ps}{%
\subsection{Vignette 11 (article
3-4-2015\_PS)}\label{vignette-11-article-3-4-2015_ps}}

Outcome: Not fully reproducible (with author involvement)

Substantial implications for the original conclusions:

R Markdown report: https://perma.cc/PGT4-GN8U

OSF/Github repository: https://osf.io/zkmgw/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.6468572.v1

Description:

In this reproducibility check we were able to reproduce the descriptive
statistics (Figure 3) but ran into difficulties reproducing some of the
inferential statistics. We contacted the authors and received assistance
that resolved some issues. Specifically, information was provided about
the identity of statistical tests, units of analysis, and normalization
procedures, that were not stated in the paper. Unfortunately, we then
lost contact with the authors and some reproducibility issues remained
that we could not resolve.

We could not reproduce two effect sizes (d). The authors reported that
they tried to reproduce these values themselves and could ``closely
reproduce the Cohen's d values''. However, they did not share the exact
values they obtained in their re-analysis. We could also not reproduce
one mean and one bound of a confidence interval. The causal locus of
these issues is unclear.

The reproducibility issues do not appear to undermine the original
conclusions. The obtained effect sizes are smaller than those reported,
but not substantially so. The mean and ci discrepancies are of low
magnitude.

\hypertarget{vignette-12-article-2-10-2014_ps}{%
\subsection{Vignette 12 (article
2-10-2014\_PS)}\label{vignette-12-article-2-10-2014_ps}}

Outcome: Reproducible (without author involvement)

Substantial implications for the original conclusions: N/A

R Markdown report: https://perma.cc/V4R9-V8JJ

OSF/Github repository: https://osf.io/2mq46/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.7548371.v1

Description:

The manuscript did not specify the way that confidence intervals were
computed and it was necessary to make an educated guess. All target
values were reproduced successfully.

\hypertarget{vignette-13-article-3-9-2014_ps}{%
\subsection{Vignette 13 (article
3-9-2014\_PS)}\label{vignette-13-article-3-9-2014_ps}}

Outcome: Not fully reproducible (with author involvement)

Substantial implications for the original conclusions: Unlikely

R Markdown report: https://perma.cc/4PFB-LHBP

OSF/Github repository: https://osf.io/b3v7h/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.4160541.v1

Description:

Whilst trying to reproduce an ANOVA we noted that two participants
(participant 25 for both the Moroccan and Spaniard samples) had missing
data for the PAST condition. However, the degrees of freedom reported in
the paper appeared to reflect that the participants \emph{were} included
in this analysis. We asked the original authors for clarification and
they informed us that there were typos in the data file they had posted
on OSF. The wrong participant ID had been added for some lines of data,
making it appear that two participants were missing data. Once the typos
were corrected, the degrees of freedom in our re-analyses matched those
reported in the paper.

Two major numerical discrepancies with p-values remained. In both cases,
the authors reported that p = .001, but the values we obtained were
lower than this. We discussed this issue with the authors and they said
that it is their policy to report p values that are \textless{}.001 as p
= .001. Unfortunately this approach guarantees inaccurate reporting.
According to the APA, researchers should ``report p values less than
.001 as p \textless{} .001'' (APA, 2009; p.~114). The reproducibility
issues we encountered do not appear to undermine the conclusions drawn
in the original article.

\hypertarget{vignette-14-article-4-1-2015_ps}{%
\subsection{Vignette 14 (article
4-1-2015\_PS)}\label{vignette-14-article-4-1-2015_ps}}

Outcome: Reproducible (with author involvement)

Substantial implications for the original conclusions: N/A

R Markdown report: https://perma.cc/R2V2-Y3HS

OSF/Github repository: https://osf.io/c6u95/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.4555168.v1

Description:

We initially could not reproduce some degrees of freedom and F-values
despite trying multiple different model specifications. We contacted the
authors for assistance and they correctly pointed out that we had missed
an important footnote (footnote 2) which read: ``The sphericity
assumption for this analysis and the corresponding analysis in
Experiments 2 and 3 was not met. We report multivariate test results, as
recommended by Maxwell and Delaney (2004), because these tests are more
optimal than correcting for sphericity. The pattern of results did not
differ depending on whether we used either a multivariate test or
sphericity correction.'' They also sent SPSS syntax and a screenshot of
the output - which demonstrated successful reproduction of the target
values.

Unfortunately at this time our team does not have access to SPSS nor the
expertise to independently implement these multivariate tests in R (the
analysis appears to be beyond our operational definition of a
`reasonably straightforward analysis'), however the SPSS syntax and
output indicates that these outcomes are reproducible.

\hypertarget{vignette-15-article-4-11-2014_ps}{%
\subsection{Vignette 15 (article
4-11-2014\_PS)}\label{vignette-15-article-4-11-2014_ps}}

Outcome: Reproducible (without author involvement)

Substantial implications for the original conclusions: N/A

R Markdown report: https://perma.cc/VA79-D42D

OSF/Github repository: https://osf.io/2kz9b/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.2298575.v1

Description:

We encountered no major difficulties reproducing all target values for
this article.

\hypertarget{vignette-16-article-5-4-2015_ps}{%
\subsection{Vignette 16 (article
5-4-2015\_PS)}\label{vignette-16-article-5-4-2015_ps}}

Outcome: Reproducible (with author involvement)

Substantial implications for the original conclusions: N/A

R Markdown report: https://perma.cc/9DEH-7YPY

OSF/Github repository: https://osf.io/bm4cg/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.1310152.v1

Description:

Although it was specified in the paper that one participant was excluded
from the analyses due to a computer malfunction, the data files provided
online did not specify which participant this referred to. Thus, prior
to communication with the author, we were not able to proceed with the
analyses. After the author identified the participant to be excluded, we
were able to reproduce all target values.

\hypertarget{vignette-17-article-6-1-2015_ps}{%
\subsection{Vignette 17 (article
6-1-2015\_PS)}\label{vignette-17-article-6-1-2015_ps}}

Outcome: Reproducible (without author involvement)

Substantial implications for the original conclusions: N/A

R Markdown report: https://perma.cc/BTS6-5MH6

OSF/Github repository: https://osf.io/f9q28/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.0405521.v1

Description:

We encountered no major difficulties reproducing all target values for
this article.

\hypertarget{vignette-18-article-6-7-2014_ps}{%
\subsection{Vignette 18 (article
6-7-2014\_PS)}\label{vignette-18-article-6-7-2014_ps}}

Outcome: Reproducible (without author involvement)

Substantial implications for the original conclusions: N/A

R Markdown report: https://perma.cc/Z8SP-CEDV

OSF/Github repository: https://osf.io/qezax/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.0223565.v1

Description:

We encountered no major difficulties reproducing all target values for
this article.

\hypertarget{vignette-19-article-7-3-2015_ps}{%
\subsection{Vignette 19 (article
7-3-2015\_PS)}\label{vignette-19-article-7-3-2015_ps}}

Outcome: Not fully reproducible (without author involvement)

Substantial implications for the original conclusions: Unknown (analysis
could not be completed)

R Markdown report: https://perma.cc/77ND-XHKN

OSF/Github repository: https://osf.io/jgp8e/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.7977267.v1

Description:

In our initial attempts, we found a number of minor numerical
discrepancies in the descriptive statistics whether we used medians or
means. We then encountered major numerical discrepancies in the
inferential statistics. There were a number of aspects of the original
analysis and data files we were unclear about. Unfortunately despite
emailing the original authors several times we have not received a
response to our questions (we have received responses saying they will
get back to us, but this has not happened \textgreater{} 20 months after
the last message). We have thus classified these issues as `insufficient
information errors'. The issues we are unclear about are outlined in
full detail in the reproducibility report. In brief, they refer to
unclear labels and codings for variables in the data file, unclear use
of means vs.~medians, unclear levels of aggregation. Despite trying out
multiple combinations of these factors based on educated guesses, we
remain unable to reproduce some values and cannot complete parts of the
analysis and it is unclear why.

\hypertarget{vignette-20-article-8-12-2014_ps}{%
\subsection{Vignette 20 (article
8-12-2014\_PS)}\label{vignette-20-article-8-12-2014_ps}}

Outcome: Not fully reproducible (with author involvement)

Substantial implications for the original conclusions: Unlikely

R Markdown report: https://perma.cc/R38G-QQKG

OSF/Github repository: https://osf.io/5hv6f/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.0413564.v1

Description:

We initially ran into a problem reproducing the exclusions reported in
the paper. We used both the SPSS and Matlab files provided by the
authors and attempted to apply the exclusions in several different ways.
As all were unsuccessful, we contacted the original authors to ask for
assistance. They created a video of themselves performing the analysis
in SPSS and reproducing the exclusion numbers reported in the paper.
Watching this video helped us to identify that we had misunderstood the
sentence in the paper that reported the exclusions, specifically: ``Only
participants for whom both d'1 and d'2 could be computed for the
analysis subset (i.e., who had non-zero counts in every cell) were
included; above chance: n = 165, at chance: n = 33.'' We had excluded
participants with missing values in any of the cells, whereas the
original analysis excluded participants with missing values in cells
relevant to the analysis subset. After being sent the video, we were
able to reproduce this part of the analysis successfully and proceeded
with the remainder of the analysis.

We were unable to reproduce two values - a t-value (0.17 reported
vs.~0.12 in our analysis) and a d-value (0.03 reported vs 0.02 in our
analysis). We contacted the authors again and they said they also could
not reproduce these values either - in their own re-analyses they obtain
the same values as we do. The reason for the differences is unknown. The
authors suggest it might be due to a change in version of the SPSS
software that was used to run the original analysis. Although they
qualify as `major numerical discrepancies according to our operational
definition, the magnitude of these differences indicates that they are
unlikely to be consequential to the authors' original conclusions.

\hypertarget{vignette-21-article-8-5-2015_ps}{%
\subsection{Vignette 21 (article
8-5-2015\_PS)}\label{vignette-21-article-8-5-2015_ps}}

Outcome: Reproducible (without author involvement)

Substantial implications for the original conclusions: N/A

R Markdown report: https://perma.cc/WN37-9T2K

OSF/Github repository: https://osf.io/uaeqw/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.2646134.v1

Description:

We encountered no major difficulties reproducing all target values for
this article.

\hypertarget{vignette-22-article-8-7-2014_ps}{%
\subsection{Vignette 22 (article
8-7-2014\_PS)}\label{vignette-22-article-8-7-2014_ps}}

Outcome: Reproducible (with author involvement)

Substantial implications for the original conclusions: N/A

R Markdown report: https://perma.cc/4779-EB7S

OSF/Github repository: https://osf.io/mxwq7/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.7903145.v1

Description:

Initially we encountered major numerical discrepancies for the analysis
of gain scores. The t-value and the effect size obtained were
considerably smaller compared to the ones reported in the paper. We
originally calculated the gain scores by averaging scores in each test
phase (pre and post) in each condition for each subject and then
subtracting the averages for the post test from the ones for the pre
test. This way of calculating scores was based on the paper's
description: ``Gain scores were calculated by subtracting each
participant's pretest score from his or her posttest score.''

We contacted the original authors and they informed us that the gain
scores were calculated in a different way, namely by subtracting pre
from post test scores for each specific lesson (e.g., plate tectonics)
and then averaging across lesson difference scores to get a mean gain
score per subject and condition. This also included excluding some
lessons for some of the participants due to absence. This way of
calculating the gain scores was not obvious based on the information in
the paper. The supplementary material also did not include any
additional information about gain scores. After correcting the way the
gain scores were calculated we were able to reproduce all target values
successfully.

\hypertarget{vignette-23-article-8-8-2014_ps}{%
\subsection{Vignette 23 (article
8-8-2014\_PS)}\label{vignette-23-article-8-8-2014_ps}}

Outcome: Reproducible (with author involvement)

Substantial implications for the original conclusions: N/A

R Markdown report: https://perma.cc/58C5-HWDQ

OSF/Github repository: https://osf.io/buegj/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.2241042.v1

Description:

We initially had trouble reproducing the principle components analysis
to generate the wise reasoning composite scores from the raw data due to
a missing variable for question 4A from Table 1 and 5 missing values in
the raw data file that were not mentioned in the paper. But after author
clarifications, these issues were resolved.

Some major numerical discrepancies remained, though their absolute
magnitude was low. The authors provided SPSS syntax and we were able to
implement this and reproduce the values reported in the paper. Despite
extensive efforts we were unable to localize the exact reason why we
could not reproduce the values in R. Our suspicion is that these might
be caused by opaque SPSS/R computational differences. Nevertheless, we
were able to reproduce all target values using a combination of R and
SPSS.

\hypertarget{vignette-24-article-9-2-2015_ps}{%
\subsection{Vignette 24 (article
9-2-2015\_PS)}\label{vignette-24-article-9-2-2015_ps}}

Outcome: Reproducible (without author involvement)

Substantial implications for the original conclusions: N/A

R Markdown report: https://perma.cc/9VCW-MXVX

OSF/Github repository: https://osf.io/cneqr/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.7685589.v1

Description:

We encountered no major difficulties reproducing all target values for
this article.

\hypertarget{vignette-25-article-9-5-2014_ps}{%
\subsection{Vignette 25 (article
9-5-2014\_PS)}\label{vignette-25-article-9-5-2014_ps}}

Outcome: Not fully reproducible (with author involvement)

Substantial implications for the original conclusions: Unknown (analysis
could not be completed)

R Markdown report: https://perma.cc/SVH6-8EN6

OSF/Github repository: https://osf.io/fam6g/

Code Ocean reproducible analysis container:
https://doi.org/10.24433/CO.8022934.v1

Description:

We were initially able to reproduce the descriptive statistics and the
three key figures. But there were four insufficient information errors
pertaining to the original analysis specifications that prevented a full
reproducibility check (see reproducibility report for details). We
contacted the original authors for assistance and they clarified some
aspects of the analyses which resolved some issues.

However, we continued to have difficulties with the section reporting
post-hoc comparisons. There were three major numerical discrepancies for
p-values. The authors report using `Sidak corrections' but it is unclear
if they were applied to the alpha threshold (correctly) or the p-values
themselves. If corrections were applied to the p values, then this could
potentially explain the differences, but the article does not identify
the family of hypotheses for which corrections were applied. So we also
cannot correct the alpha threshold. We have decided to report the major
numerical discrepancies for the p values, but also record an
insufficient information error as we cannot determine whether these were
decision errors or not. Additionally, for one degree of freedom, we
obtained 4 whereas the article reports 3. This may be a typo but the
cause is not clear.

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-kidwell2016}{}%
Kidwell, M. C., Lazarevi\a'c, L. B., Baranski, E., Hardwicke, T. E.,
Piechowski, S., Falkenberg, L.-S., Kennett, C., Slowik, A., Sonnleitner,
C., Hess-Holden, C., Errington, T. M., Fiedler, S., \& Nosek, B. A.
(2016). Badges to Acknowledge Open Practices: A Simple, Low-Cost,
Effective Method for Increasing Transparency. \emph{PLOS Biology},
\emph{14}(5), e1002456.
\url{https://doi.org/10.1371/journal.pbio.1002456}
\end{appendix}